---
title: "Lotus MCMC Line Painting Demo"
author: "David Hodgson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lotus MCMC Line Painting Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)
```

# Lotus MCMC Line Painting Demo

This vignette demonstrates the MCMC line painting algorithm applied to a beautiful lotus flower image. The algorithm will run for 100,000 iterations to create a high-quality artistic interpretation.

## Load the Package

```{r load-package}
# Load the mcmcPainter package
library(mcmcPainter)
```

## Load and Display the Target Image

```{r load-image}
# Load the lotus image
image_path <- "../inst/extdata/lotus.png"

# Display the target image
if (file.exists(image_path)) {
  # Load and display the image
  img <- png::readPNG(image_path)
  plot.new()
  rasterImage(img, 0, 0, 1, 1)
  title("Target Lotus Image")
} else {
  cat("Image not found at:", image_path, "\n")
}
```

## Run MCMC Line Painting

```{r run-mcmc, eval=FALSE}
# Run the MCMC line painting algorithm
# Note: This is a long-running process (several hours for 100,000 iterations)
res <- run_line_painter(
  image_path = image_path,
  max_dimension = 1200,
  iters = 100000,
  out_dir = "../inst/results/lotus_line_painting_100k",
  seed = 42,
  auto_config = TRUE,
  save_every = 5000,
  verbose = TRUE
)
```

## Create Triptych Visualization

```{r create-triptych, eval=FALSE}
# Create a triptych showing the progression
triptych_path <- create_triptych(
  image_path = image_path,
  result = res,
  out_path = "../inst/results/lotus_line_triptych.png",
  max_dimension = 1200
)
```

## Display Results

```{r display-results, eval=FALSE}
# Display the final triptych
if (file.exists("../inst/results/lotus_line_triptych.png")) {
  triptych <- png::readPNG("../inst/results/lotus_line_triptych.png")
  plot.new()
  rasterImage(triptych, 0, 0, 1, 1)
  title("Lotus MCMC Line Painting Triptych")
} else {
  cat("Triptych not yet created. Run the MCMC algorithm first.\n")
}
```

## Algorithm Details

The MCMC line painting algorithm uses the following approach:

1. **Initialization**: Start with a white canvas
2. **Iterative Process**: For each iteration:
   - **Birth Move**: Add new lines based on image residuals
   - **Death Move**: Remove existing lines
   - **Jitter Move**: Perturb line parameters (position, color, opacity, thickness)
   - **Swap Move**: Reorder line rendering for better composition
3. **Acceptance**: Accept or reject moves based on likelihood improvement
4. **Adaptive Temperature**: Gradually increase acceptance rate over time

## Performance Notes

- **Iterations**: 100,000 steps for high-quality results
- **Save Frequency**: Every 5,000 iterations
- **Canvas Size**: Automatically scaled to 1200px maximum dimension
- **Expected Runtime**: Several hours on modern hardware

## Results Interpretation

The final triptych will show:
- **Left Panel**: Initial white canvas
- **Middle Panel**: Best MCMC result after 100,000 iterations
- **Right Panel**: Original target image

This allows for easy comparison of the algorithm's performance and artistic interpretation quality.
