---
title: "Me MCMC Line Painting Demo"
author: "David Hodgson"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
    code_folding: show
  pdf_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Me MCMC Line Painting Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 8,
  fig.align = "center"
)

# Load required packages
if (!require(magick)) install.packages("magick")
if (!require(png)) install.packages("png")
if (!require(Rcpp)) install.packages("Rcpp")
```

# Me MCMC Line Painting Demo

This vignette demonstrates the complete workflow for generating line paintings using the `mcmcPainter` package, featuring the personal `me.png` image with 100,000 MCMC iterations for maximum detail and quality.

## ðŸ‘¤ **Image Analysis**

First, let's analyze the me image to understand its characteristics:

```{r image_analysis}
# Load package functions
source("../R/mcmcPainter.R")
source("../R/mcmc_core.R") 
source("../R/utilities.R")

# Analyze the me image
image_path <- "../inst/extdata/me.png"
cat("Analyzing image:", basename(image_path), "\n")
cat("=====================================\n")

img_info <- get_image_info(image_path)
cat("File:", basename(image_path), "\n")
cat("Original dimensions:", img_info$width, "x", img_info$height, "pixels\n")
cat("File size:", round(img_info$file_size / 1024, 1), "KB\n")
cat("PNG verification:", ifelse(img_info$is_true_png, "âœ“ True PNG", "âœ— Not a true PNG"), "\n")
cat("Aspect ratio:", round(img_info$width / img_info$height, 3), "\n")
cat("Total pixels:", format(img_info$width * img_info$height, big.mark = ","), "\n\n")

# Display the original image
library(magick)
me_img <- image_read(image_path)
image_info <- image_info(me_img)
cat("Magick image info:\n")
cat("Dimensions:", image_info$width, "x", image_info$height, "\n")
cat("Format:", image_info$format, "\n")
cat("Colorspace:", image_info$colorspace, "\n")
cat("Filesize:", round(image_info$filesize / 1024, 1), "KB\n\n")
```

## ðŸŽ¯ **MCMC Configuration**

Now let's configure the MCMC parameters for optimal performance on the me image:

```{r mcmc_config}
# High-quality configuration for 100K steps
cat("High-Quality MCMC Configuration:\n")
cat("================================\n")
cat("Target iterations: 100,000\n")
cat("Save frequency: Every 5,000 iterations\n")
cat("Max dimension: 1200 (high resolution)\n")
cat("Expected runtime: 3-5 hours (depending on system)\n\n")

# Auto-configure MCMC parameters
config <- auto_configure_mcmc(
  image_path = image_path,
  max_dimension = 1200,        # High resolution
  target_iterations = 100000    # 100K iterations
)

cat("Optimized MCMC Parameters:\n")
cat("==========================\n")
cat("Original dimensions:", config$original_width, "x", config$original_height, "\n")
cat("Scaled dimensions:", config$scaled_width, "x", config$scaled_height, "\n")
cat("Scaling factor:", round(config$scale_factor, 3), "\n")
cat("Adjusted iterations:", format(config$iterations, big.mark = ","), "\n")
cat("Save frequency:", format(config$save_every, big.mark = ","), "\n")
cat("PNG verification:", ifelse(config$is_true_png, "âœ“ True PNG", "âœ— Not a true PNG"), "\n\n")

# Calculate complexity metrics
original_pixels <- config$original_width * config$original_height
scaled_pixels <- config$scaled_width * config$scaled_height
complexity_reduction <- (1 - scaled_pixels / original_pixels) * 100

cat("Complexity Analysis:\n")
cat("===================\n")
cat("Original pixels:", format(original_pixels, big.mark = ","), "\n")
cat("Scaled pixels:", format(scaled_pixels, big.mark = ","), "\n")
cat("Complexity reduction:", round(complexity_reduction, 1), "%\n")
cat("Memory usage:", round(scaled_pixels * 3 * 8 / 1024 / 1024, 1), "MB\n\n")
```

## ðŸš€ **Running the MCMC Algorithm**

Now let's run the MCMC algorithm to generate the line painting. This will take several hours for 100K iterations:

```{r mcmc_run, eval=FALSE}
# Compile the C++ code for performance
cat("Compiling C++ code...\n")
Rcpp::sourceCpp("../src/mcmc_painter_cpp.cpp")
cat("C++ code compiled successfully!\n\n")

# Run MCMC with high-quality settings
cat("Starting high-quality MCMC run for me...\n")
cat("This will take several hours. Progress will be saved every 5,000 iterations.\n\n")

res <- run_line_painter(
  image_path = image_path,
  max_dimension = 1200,        # High resolution
  iters = 100000,              # 100K iterations
  out_dir = "../inst/results/me_100k_high_quality",
  seed = 42,
  auto_config = TRUE,          # Enable auto-configuration
  verbose = TRUE
)

cat("\nðŸŽ‰ MCMC completed successfully!\n")
cat("===============================\n")
cat("Final number of lines:", length(res$lines), "\n")
cat("Best iteration:", res$best$iter, "\n")
cat("Best SSE:", round(res$best$sse, 2), "\n")
```

## ðŸ“Š **Results Analysis**

After the MCMC completes, let's analyze the results:

```{r results_analysis, eval=FALSE}
# Load the final image with the same dimensions used in MCMC
target_img <- load_image_rgb(image_path, 
                            out_w = dim(res$canvas)[2], 
                            out_h = dim(res$canvas)[1])

# Create default white canvas
H <- dim(target_img)[1]
W <- dim(target_img)[2]
default_canvas <- array(1, dim = c(H, W, 3))

# Calculate performance metrics
sse <- sum((target_img - res$best$canvas)^2)
mse <- sse / length(target_img)
psnr <- 20 * log10(1 / sqrt(mse))

cat("Performance Metrics:\n")
cat("==================\n")
cat("Best SSE:", round(sse, 2), "\n")
cat("Best MSE:", round(mse, 6), "\n")
cat("Best PSNR:", round(psnr, 2), "dB\n")
cat("Number of lines:", length(res$lines), "\n")
cat("Final dimensions:", W, "x", H, "\n\n")

# Display optimization summary
cat("Optimization Summary:\n")
cat("====================\n")
cat("Original dimensions:", config$original_width, "x", config$original_height, "\n")
cat("Final dimensions:", W, "x", H, "\n")
cat("Total iterations:", 100000, "\n")
cat("Save frequency:", 5000, "\n")
cat("PNG verification:", ifelse(config$is_true_png, "âœ“ True PNG", "âœ— Not a true PNG"), "\n")
```

## ðŸŽ¨ **Triptych Visualization**

Create a beautiful triptych showing the progression from white canvas to final artwork:

```{r triptych_creation, eval=FALSE}
# Create triptych
create_triptych(
  default_canvas = default_canvas,
  best_canvas = res$best$canvas,
  target_img = target_img,
  titles = c("Default (White Canvas)", "Best MCMC Result (100K steps)", "True Me Image"),
  main_title = "Me Image: 100K MCMC Progression"
)

# Save triptych in multiple formats
cat("Saving triptych...\n")
output_dir <- "../inst/results/me_100k_high_quality"

# Save as PDF
pdf_path <- file.path(output_dir, "me_100k_triptych.pdf")
save_triptych(
  default_canvas = default_canvas,
  best_canvas = res$best$canvas,
  target_img = target_img,
  output_path = pdf_path,
  width = 18, height = 7,  # Larger for high-quality output
  titles = c("Default (White Canvas)", "Best MCMC Result (100K steps)", "True Me Image"),
  main_title = "Me Image: 100K MCMC Progression",
  format = "pdf"
)

# Save as PNG
png_path <- file.path(output_dir, "me_100k_triptych.png")
save_triptych(
  default_canvas = default_canvas,
  best_canvas = res$best$canvas,
  target_img = target_img,
  output_path = png_path,
  width = 18, height = 7,  # Larger for high-quality output
  titles = c("Default (White Canvas)", "Best MCMC Result (100K steps)", "True Me Image"),
  main_title = "Me Image: 100K MCMC Progression",
  format = "png"
)

cat("Triptych saved successfully!\n")
cat("Files created:\n")
cat("- PDF:", pdf_path, "\n")
cat("- PNG:", png_path, "\n")
```

## ðŸ“ **Output Files**

The MCMC run will generate several output files:

```{r output_files, eval=FALSE}
# List all saved iterations
cat("Saved Iterations:\n")
cat("=================\n")
iter_files <- list.files(output_dir, pattern = "iter_.*\\.png", full.names = FALSE)
iter_files <- sort(iter_files)
for (file in iter_files) {
  cat("-", file, "\n")
}

cat("\nFinal Output Files:\n")
cat("==================\n")
cat("- Best canvas: best_canvas.png\n")
cat("- Final lines: final_lines.RData\n")
cat("- Triptych PDF: me_100k_triptych.pdf\n")
cat("- Triptych PNG: me_100k_triptych.png\n")
cat("- Progress log: mcmc_progress.log\n")
```

## ðŸ” **Progression Analysis**

Let's examine how the MCMC progresses over the 100K iterations:

```{r progression_analysis, eval=FALSE}
# Load saved iterations for analysis
library(png)
library(grid)

# Function to load and display progression
show_progression <- function(iterations = c(0, 25000, 50000, 75000, 100000)) {
  par(mfrow = c(2, 3), mar = c(2, 2, 2, 1))
  
  for (iter in iterations) {
    if (iter == 0) {
      # Show white canvas
      img <- array(1, dim = c(H, W, 3))
      title <- "Iteration 0 (White)"
    } else {
      # Load saved iteration
      file_path <- file.path(output_dir, paste0("iter_", sprintf("%06d", iter), ".png"))
      if (file.exists(file_path)) {
        img <- readPNG(file_path)
        title <- paste("Iteration", format(iter, big.mark = ","))
      } else {
        next
      }
    }
    
    plot.new()
    rasterImage(img, 0, 0, 1, 1)
    title(title, cex.main = 0.8)
  }
  
  # Show target image
  plot.new()
  rasterImage(target_img, 0, 0, 1, 1)
  title("Target Image", cex.main = 0.8)
  
  par(mfrow = c(1, 1))
}

# Display progression (if files exist)
if (dir.exists(output_dir)) {
  show_progression()
}
```

## ðŸ“ˆ **Performance Comparison**

Compare the me results with other images:

```{r performance_comparison, eval=FALSE}
# Performance comparison table
comparison_data <- data.frame(
  Image = c("Me", "Butterfly", "Octopus", "Leaf", "Iamami"),
  Dimensions = c(
    paste(config$scaled_width, "x", config$scaled_height),
    "2068 x 2091", 
    "964 x 900",
    "800 x 1422",
    "788 x 605"
  ),
  Iterations = c(100000, 100000, 100000, 20000, 100000),
  File_Size_KB = c(
    round(config$file_size / 1024, 1),
    4356,
    380,
    132,
    336
  ),
  Complexity = c(
    "High (3.2MB)",
    "Very High (4.3MB)",
    "Medium (380KB)",
    "Medium (132KB)", 
    "Medium (336KB)"
  )
)

knitr::kable(comparison_data, 
             caption = "Performance Comparison Across Images",
             align = c('l', 'c', 'c', 'c', 'l'))
```

## ðŸŽ¯ **Key Insights**

### **Why 100K Iterations for Me?**

1. **High Complexity**: The me image likely has intricate facial features and details
2. **Large File Size**: 3.2MB indicates rich texture and color variation
3. **Fine Details**: Facial features, hair, and expressions require precise line placement
4. **Color Gradients**: Skin tones and lighting need many thin, overlapping strokes

### **Expected Results**

- **Line Count**: 1000-2000+ lines for full detail
- **Quality**: Near-photorealistic line painting
- **Convergence**: Stable improvement through 100K steps
- **Memory**: Efficient bounding box updates for performance

## ðŸš€ **Running the Full Demo**

To run the complete 100K MCMC demo:

```r
# Option 1: Run the vignette
rmarkdown::render("vignettes/me_mcmc_demo.Rmd")

# Option 2: Run directly in R
source("vignettes/me_mcmc_demo.Rmd")

# Option 3: Use the standalone script
source("create_me_triptych.R")
```

## ðŸ“š **Further Reading**

- **Package Documentation**: `?mcmcPainter`
- **Core Functions**: `?run_line_painter`, `?create_triptych`
- **Utilities**: `?get_image_info`, `?auto_configure_mcmc`
- **Vignettes**: See other image demos for comparison

---

**Note**: The 100K MCMC run will take 3-5 hours depending on your system. The results will show the full artistic potential of the algorithm on complex, personal images like me! ðŸ‘¤âœ¨
